\documentclass[12pt,a4paper]{article}
\usepackage{t1enc}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\topmargin -1cm
\oddsidemargin 0cm

\textwidth  16cm
\textheight 24cm
\pagestyle{empty}

\begin{document}

%%%%%%%% title %%%%%%%%%%
\title{{\bf EmBOX} networking subsystem: refactoring raw sockets - present state(draft.0.1)}
\date{February 3, 2012}
\author{Timur Abdukadyrov}

\maketitle
%%%%%%%%%%%%%%%%%%%%%%%%

\newpage
\tableofcontents

%%%%%%%%%%% document description %%%%%%%%%%%%
\newpage

\section{Document description}
\label{sec:description}
Current version of this document is 0.1(draft).

The purpose of this document is to systemize the structure of EMBOX' sockets
networking subsystem(further ENSS), particulary raw sockets and help in refactoring
current version of that subsystem to consitute more logic and finished module. In
this text the description of files, structures and methods of socket networking
subsystem is given along with it's common logic.

Once the common structure of refactored ENSS is achieved document's version should
evaluate to 0.2. On completion of refactoring this document should reflect the
changes and version should evaluate to 0.3 and become description of ENSS.

\newpage
\section{TODO}
This document still lacks the description of shared socket methods and ip protocol
family description.
\section{Refactoring}
\begin{itemize}
  \item More logic file tree structure in {\it net/} and particulary in
    {\it net/ipv4}
  \item Develop a generalized shared IP methods library to have a flexible
    instrument for IP functioning.
  \item Raw socket can fetch a supposed protocol, which can be omited (Linux
    style). And it should realize respected processing of headers. For example
    if a raw socket is a UDP raw socket than, UDP headers can be left to be
    processed by kernel itself and we are only interested in the following
    packet contets. In case we omit the protocol of raw socket we should take
    care of the whole packet contents begining with IP header(classic variant).
    Previously developed shared methods library should be the sub-layer for
    this one.
  \item Other higher level protocols now could be realized via only raw
    sockets with a defined set of preferencies and propetries.
\end{itemize}
\newpage

\section{socket.c}
\label{sec:socket_c_socket.c}
\subsection{Functions}
\label{sec:socket_c_socket_functions}
{\it
\begin{itemize}
  \item this\_read
  \item this\_write
  \item id2sock
  \item socket
  \item connect
  \item bind
  \item listen
  \item accept
  \item sendto\_sock
  \item sendto
  \item check\_icmp\_err
  \item recvfrom\_sock
  \item recvfrom
  \item socket\_close
\end{itemize}
}

\subsubsection{this\_read}
\label{sec:socket_c_this_read}
\begin{verbatim}
static ssize_t this_read(int fd, const void *buf, size_t nbyte)
\end{verbatim}
This is a {\it task\_res\_ops} read member. {\it task\_res\_ops} is defined
in {\it include/kernel/task\_resources.h}.


\subsubsection{this\_write}
\label{sec:socket_c_this_write}
\begin{verbatim}
static ssize_t this_write(int fd, const void *buf, size_t nbyte)
\end{verbatim}
This is a {\it task\_res\_ops} write member.  {\it task\_res\_ops} is defined
in {\it include/kernel/task\_resources.h}.

\subsubsection{idx2sock}
\label{sec:socket_c_idx2sock}
\begin{verbatim}
static struct socket *idx2sock(int fd)
\end{verbatim}
Socket descriptor number to {\it socket} conversion.

\subsubsection{socket}
\label{sec:socket_c_socket}
\begin{verbatim}
int socket(int domain, int type, int protocol)
\end{verbatim}
Create a socket.

{\bf Returns} error code if any. Else returns
{\it task\_res\_idx\_alloc} result, which is defined in
{\it include/kernel/task\_resources.h} and {\it kernel/task/index\_desc.c}.

{\bf Is a} wrap for {\it kernel\_socket\_create} function plus trivial checks.

\subsubsection{connect}
\label{sec:socket_c_connectn}
\begin{verbatim}
int connect(int sockfd, const struct sockaddr *daddr, socklen_t daddrlen)
\end{verbatim}
Connect to destination address {\it daddr}.

{\bf Returns} socket connection result error code

{\bf Is a} wrap for {\it kernel\_socket\_connect} function  plus trivial checks.

\subsubsection{bind}
\label{sec:socket_c_bind}
\begin{verbatim}
int bind(int sockfd, const struct sockaddr *addr, socklen_t addrlen)
\end{verbatim}
Bind socket to address {\it addr}.

{\bf Returns} socket binding result error code.

{\bf Is a} wrap for {\it kernel\_socket\_bind} function  plus trivial checks.

\subsubsection{listen}
\label{sec:socket_c_listen}
\begin{verbatim}
int listen(int sockfd, int backlog)
\end{verbatim}
Begin listenning on a socket.

{\bf Returns} socket listening result error code.

{\bf Is a} wrap for {\it kernel\_socket\_listen} function  plus trivial checks.


\subsubsection{accept}
\label{sec:socket_c_accept}
\begin{verbatim}
int accept(int sockfd, struct sockaddr *addr, socklen_t *addrlen)
\end{verbatim}
Accept incoming connection.

{\bf Returns} socket acceptance result error code if any. Else returns socket
descriptor.

{\bf Is a} wrap for {\it kernel\_socket\_accept} function  plus trivial checks.


\subsubsection{sendto\_sock}
\label{sec:socket_c_sendto_sock}
\begin{verbatim}
static size_t sendto_sock(struct socket *sock, const void *buf, size_t len,
       int flags, const struct sockaddr *daddr, socklen_t daddrlen)
\end{verbatim}
Structurizes the input buffer and fills out {\it inet} (of {\bf \it inet\_sock}
type). The {\it sport} member of {\it inet} check and filling if needed.
Overwrites {\it sock->sk->is\_ready} and {\it sock->sk\_err}. Then call to
{\it kernel\_socket\_sendmsg}. Uses also {\it socket\_set\_port\_type()} and
{\it socket\_get\_free\_port()} functions from net/socket/port.c.

{\bf Returns} Error code if any. Else returns {\it len} input argument
unmodified.

\subsubsection{sendto}
\label{sec:socket_c_sendto}
\begin{verbatim}
ssize_t sendto(int sockfd, const void *buf, size_t len, int flags,
               const struct sockaddr *daddr, socklen_t daddrlen)
\end{verbatim}
Send datagram to socket, using socket descriptor. User-level interface(?)

{\bf Returns} {\it sendto\_sock} result.

{\bf Is a } wrap for {\it sendto\_sock} function.

\subsubsection{check\_icmp\_err}
\label{sec:socket_c_check_icmp_err}
\begin{verbatim}
int check_icmp_err(int sockfd)
\end{verbatim}
Check for errors from icmp handler in icmp module. Get errors, set
{\it sock->sk->sk\_err} = -1.

{\bf Returns} last error.

\subsubsection{recvfrom\_sock}
\label{sec:socket_c_recvfrom_sock}
\begin{verbatim}
static ssize_t recvfrom_sock(struct socket *sock, void *buf, size_t len,
       int flags, struct sockaddr *daddr, socklen_t *daddrlen)
\end{verbatim}
Receive data from socket {\it sock}. Prepare {\it iovec} and {\it msghdr}
structures and call {\it kernel\_socket\_recvmsg}.

{\bf Returns } {\it res}, which is the return value of
{\it kernel\_socket\_recvmsg()}.

\subsubsection{recvfrom}
\label{sec:socket_c_recvfrom}
\begin{verbatim}
ssize_t recvfrom(int sockfd, void *buf, size_t len, int flags,
        struct sockaddr *daddr, socklen_t *daddrlen)
\end{verbatim}
{\bf Is a} wrap for {\it recvfrom\_sock()}.

{\bf Returns} what {\it recvfrom\_sock()} returns.

\subsubsection{socket\_close}
\label{sec:socket_c_socket_close}
\begin{verbatim}
int socket_close(int sockfd)
\end{verbatim}
Find socket matching socket descriptor number and try to close it.

{\bf Returns} result of {\it kernel\_socket\_release()} if socket descriptor
was ok. Else returns {\it -EBADF} value.

%%%%%%%%%%%% socket.c structures %%%%%%%%%
\subsection{Structures}
\subsubsection{iovec}
\begin{verbatim}
struct iovec {
  void *iov_base;
  size_t iov_len;
};
\end{verbatim}
Input/output vector.

{\bf Is used} in {\it msghdr} structure.

\subsubsection{msghdr}
\label{sec:socket_c_msghdr}
\begin{verbatim}
struct msghdr {
  void *msg_name;       /* Socket name */
  int msg_namelen;      /* Length of name */
  struct iovec *msg_iov; /* Data blocks */
#if 0
  __kernel_size_t msg_iovlen;  /* Number of blocks    */
#endif
  void *msg_control;    /* Per protocol magic (eg BSD file descriptor passing) */
#if 0
  __kernel_size_t  msg_controllen;  /* Length of cmsg list */
#endif
  unsigned int msg_flags;
};
\end{verbatim}
{\bf Is used} in {\it *\_recvmsg} and {\it *\_sendmsg} functions in
\begin{itemize}
\item   net/ipv4/af\_inet.c
\item   net/ipv4/tcp.c
\item   net/ipv4/icmp.c
\item   net/ipv4/raw\_sock.c
\item   net/ipv4/udp.c
\item   net/core/sock.c
\item   net/socket/socket.c
\item   net/socket/kernel\_socket.c
\end{itemize}

\subsubsection{sockaddr}
\label{sec:socket_c_sockaddr}
\begin{verbatim}
typedef struct sockaddr {
	sa_family_t sa_family; /* Address family, AF_xxx */
	char sa_data[14];      /* 14 bytes of protocol address */
} sockaddr_t;
\end{verbatim}
Matches {\it osockaddr} in BSD. {\it sa\_family} data type is alias to
{\it unsigned short}.

It is often overcasted as {\it sockaddr\_in}. It's definition is as follows
\begin{verbatim}
struct sockaddr_in {
  short            sin_family;   /* e.g. AF_INET */
  unsigned short   sin_port;     /* e.g. htons(3490) */
  unsigned short   port_type;
  struct in_addr   sin_addr;     /* see struct in_addr, below */
  char             sin_zero[8];  /* zero this if you want to */
};
\end{verbatim}
It is defined in {\it include/net/in.h}.

%%%%%%%%%%%% kernel_socket.c %%%%%%%%%%%%%
\newpage
\section{kernel\_socket.c}
\label{sec:kernel_socket_kernel_socket.c}

\subsection{Functions}
\label{sec:kernel_socket_kernel_socket_functions}

\subsubsection{kernel\_socket\_create}
\label{sec:kernel_socket_kernel_socket_create}
\begin{verbatim}
int kernel_socket_create(int family, int type, int protocol,
                         struct socket **psock)
\end{verbatim}
Create a socket. Calls {\it socket\_alloc()} from
{\it net/socket/socket\_allocator.c}. And then calls socket type native method
{\it create()} from {\it net\_proto\_family} structure of this family. If
nothing happened free allocated socket via {\it socket\_free()} same from
{\it socket\_allocator.c} (also here stays TODO about socket permissions).

{\bf Returns} {\it -EINVAL} if checks regarding type and family don't pass.
{\it -ENOMEM} if {\it socket\_alloc()} couldn't allocate memory. Result returned
by native {\it create()} method is returned elsewise.

\subsubsection{kernel\_socket\_release}
\label{sec:kernel_socket_kernel_socket_release}
\begin{verbatim}
int kernel_socket_release(struct socket *sock)
\end{verbatim}
Release {\it {\bf socket}} structure bound to this socket. Calls native
{\it release()} method of this PF and then calls {\it socket\_free()} to
deallocate space.

{\bf Returns} {\it ENOERR} on success. Else returns error code returned by
{\it release()} method.

\subsubsection{kernel\_socket\_bind}
\label{sec:kernel_socket_kernel_socket_bind}
\begin{verbatim}
int kernel_socket_bind(struct socket *sock, const struct sockaddr *addr,
			socklen_t addrlen)
\end{verbatim}
{\bf Is a} wrap for this' PF native {\it bind()} method.

{\bf Returns} result of the method.

\subsubsection{kernel\_socket\_listen}
\label{sec:kernel_socket_kernel_socket_listen}
\begin{verbatim}
int kernel_socket_listen(struct socket *sock, int backlog)
\end{verbatim}
{\bf Is a} wrap for this' PF native {\it listen()} method.

{\bf Returns} result of the method.

\subsubsection{kernel\_socket\_accept}
\label{sec:kernel_socket_kernel_socket_accept}
\begin{verbatim}
int kernel_socket_accept(struct socket *sock, struct sockaddr *addr,
                         socklen_t *addrlen)
\end{verbatim}
{\bf Is a} wrap for this' PF native {\it accept()} method.

{\bf Returns} result of the method.


\subsubsection{kernel\_socket\_connect}
\label{sec:kernel_socket_kernel_socket_connect}
\begin{verbatim}
int kernel_socket_connect(struct socket *sock, const struct sockaddr *addr,
                          socklen_t addrlen, int flags)
\end{verbatim}
{\bf Is a} wrap for this' PF native {\it connect()} method.

{\bf Returns} result of the method.


\subsubsection{kernel\_socket\_getsockname}
\label{sec:kernel_socket_kernel_socket_getsockname}
\begin{verbatim}
int kernel_socket_getsockname(struct socket *sock,
                              struct sockaddr *addr, int *addrlen)
\end{verbatim}
{\bf Is a} wrap for this' PF native {\it getname()} method.

{\bf Returns} result of the method.


\subsubsection{kernel\_socket\_getpeername}
\label{sec:kernel_socket_kernel_socket_getpeername}
\begin{verbatim}
int kernel_socket_getpeername(struct socket *sock,
                              struct sockaddr *addr, int *addrlen)
\end{verbatim}
{\bf Is a} wrap for this' PF native {\it getname()} method.

{\bf Returns} result of the method.


\subsubsection{kernel\_socket\_getsockopt}
\label{sec:kernel_socket_kernel_socket_getsockopt}
\begin{verbatim}
int kernel_socket_getsockopt(struct socket *sock, int level, int optname,
                             char *optval, int *optlen)
\end{verbatim}
{\bf Is a} wrap for this' PF native {\it getsockopt()} method.

{\bf Returns} result of the method.


\subsubsection{kernel\_socket\_setsockopt}
\label{sec:kernel_socket_kernel_socket_setsockopt}
\begin{verbatim}
int kernel_socket_setsockopt(struct socket *sock, int level, int optname,
                             char *optval, int optlen)
\end{verbatim}
{\bf Is a} wrap for this' PF native {\it setsockopt()} method.

{\bf Returns} result of the method.


\subsubsection{kernel\_socket\_sendmsg}
\label{sec:kernel_socket_kernel_socket_sendmsg}
\begin{verbatim}
int kernel_socket_sendmsg(struct kiocb *iocb, struct socket *sock,
                          struct msghdr *m, size_t total_len)
\end{verbatim}
{\bf Is a} wrap for this' PF native {\it sendmsg()} method.

{\bf Returns} result of the method.


\subsubsection{kernel\_socket\_recvmsg}
\label{sec:kernel_socket_kernel_socket_recvmsg}
\begin{verbatim}
int kernel_socket_recvmsg(struct kiocb *iocb, struct socket *sock,
                          struct msghdr *m, size_t total_len, int flags)
\end{verbatim}
{\bf Is a} wrap for this' PF native {\it recvmsg()} method.

{\bf Returns} result of the method.

%%%%%%%%%%%%% EMBOX_NET_* %%%%%%%%%%%%%%%%%%%%
\newpage
\section {EMBOX\_NET\_PACK and EMBOX\_NET\_SOCK}
\label{sec:embox_net_pack_and_sock}
\subsection{EMBOX\_NET\_PACK}
\label{sec:embox_net_pack}
EMBOX\_NET\_PACK stands for
\begin{verbatim}
#define EMBOX_NET_PACK(_type, _func, _init)
  int _func(sk_buff_t *skb,
        net_device_t *dev,
        packet_type_t *pt,
        net_device_t *orig_dev);
  static int _init(void);
  static packet_type_t _##_type = {
      .type = _type,
      .func = _func,
      .init = _init
    };
    __EMBOX_NET_PACK(_##_type)
\end{verbatim}
and is defined in {\it include/framework/net/pack/self.h}. So for example
\begin{verbatim}
EMBOX_NET_PACK(ETH_P_IP, ip_rcv, inet_init);
\end{verbatim}
from {\it net/ipv4/af\_inet.c} after first step preprocessing becomes like
\begin{verbatim}
  int ip_rcv(sk_buff_t *skb,
        net_device_t *dev,
        packet_type_t *pt,
        net_device_t *orig_dev);
  static int inet_init(void);
  static packet_type_t _ETH_P_IP = {
      .type = ETH_P_IP,
      .func = ip_rcv,
      .init = inet_init
    };
    __EMBOX_NET_PACK(_ETH_P_IP)
\end{verbatim}
Underscore version {\it \_\_EMBOX\_NET\_PACK} stands for
\begin{verbatim}
#define __EMBOX_NET_PACK(_packet)
  extern const struct mod_ops __net_pack_mod_ops;
  extern const struct net_pack __net_pack_registry[];
  ARRAY_SPREAD_ADD_NAMED(__net_pack_registry, __net##_packet, {
    .netpack = &_packet,
    .mod = &mod_self
  });
  MOD_INFO_BIND(&__net_pack_mod_ops, &__net##_packet)
\end{verbatim}
{\it ARRAY\_SPREAD\_ADD\_NAMED} in it's turn adds to
{\it \_\_net\_pack\_registry} array entry regarding new packet family(i.e. in
our particular case {\it \_\_netETH\_P\_IP}) and {\it MOD\_INFO\_BIND}
binds native operations for that new packet family.

 The argument {\it \_type } in {\it EMBOX\_NET\_PACK} is the type of protocol field of
network packet header. After binding all data needed to process IP protocol
(i.e. after {\it EMBOX\_NET\_PACK(ETH\_P\_IP, ip\_rcv, inet\_init);}) EMBOX now
knows that each time a packet containing {\it ETH\_P\_IP}(which is standard
0x0800 for IP) protocol type in header arrives it should pass this packet to
{\it af\_inet} module.

%% EMBOX_NET_SOCK macro

\subsection{EMBOX\_NET\_SOCK}
\label{sec:embox_net_sock}

{\it EMBOX\_NET\_SOCK} stands for
\begin{verbatim}
#define EMBOX_NET_SOCK(_type, _protocol, _prot, _ops, _no_check,
            net_proto_family)
  static inet_protosw_t _##_type = {
      .type = _type,
      .protocol = _protocol,
      .prot = &_prot,
      .ops = &_ops,
      .no_check = _no_check
    };
    __EMBOX_NET_SOCK(_##_type, net_proto_family)
\end{verbatim}
and is defined in {\it include/framework/net/sock/self.h}.

Just like {\it EMBOX\_NET\_PACK} macro {\it EMBOX\_NET\_SOCK} registers a new
socket type in system.

To be complete here is what underscore version {\it \_\_EMBOX\_NET\_SOCK} does
\begin{verbatim}
#define __EMBOX_NET_SOCK(_sock, _net_proto_family)
  extern const struct mod_ops __net_sock_mod_ops;
  extern const struct net_sock __net_sock_registry[];
  ARRAY_SPREAD_ADD_NAMED(__net_sock_registry, __net##_sock, {
      .net_proto_family = _net_proto_family,
      .netsock = &_sock,
      .mod = &mod_self
    });
  MOD_INFO_BIND(&__net_sock_mod_ops, &__net##_sock )
\end{verbatim}

%%%%%%%%%%%%% af_inet.c %%%%%%%%%%%%%%%%%%%%%%
\newpage

\section{af\_inet.c}
\label{sec:af_inet_c}
Sockets of AF\_INET family.

\subsection{Functions}

\subsubsection{inet\_create}
\label{sec:embox_net_inet_create}
\begin{verbatim}
static int inet_create(struct socket *sock, int protocol)
\end{verbatim}
Create a socket. Is bound to {\it inet\_dgram\_ops.create} method.

Looks through existing socket types in EMBOX socket type array. Once the socket
type is found check protocol type.

If appropriate socket type is found allocate space for socket and initialize
structure fields. Bind {\it sock} entitiy to {\it sk} entity. And call native
{\it init()} method.

{\bf Returns} {\it -ENOENT} in case no appropriate socket was found or if
couldn't allocate space for {\it sk}. Else 0.

Various {\bf TODO}s regarding logic.

\subsubsection{inet\_release}
\label{sec:embox_net_release}
\begin{verbatim}
int inet_release(struct socket *sock)
\end{verbatim}
Release socket. Unlock ports used and call {\it close()} method of
{\it sk\_prot}. Deallocation of {\it sk} should be conducted in {\it close()}
method.

{\bf Returns} {\it -EINVAL} in case there's no {\it sk}(==NULL). {\it ENOERR}
elsewise.

\subsubsection{inet\_bind}
\label{sec:embox_net_inet_bind}
\begin{verbatim}
int inet_bind(struct socket *sock, struct sockaddr *addr, int addr_len)
\end{verbatim}
Bind socket to adress/port specified in struct {\it addr} using native protocol
{\it bind()} method.

{\bf Returns} Error code of {\it bind()} method, if errors occured. Returns
{\it -EBUSY} if the source port in {\it addr} is busy. Returns {\it NOERR}
if everything went fine.

\subsubsection{inet\_dgram\_connect}
\label{sec:embox_net_inet_dgram_connect}
\begin{verbatim}
int inet_dgram_connect(struct socket *sock, struct sockaddr * addr,
                       int addr_len, int flags)
\end{verbatim}
Datagram-connect. Find a free port, lock it for usage and call native
{\it connect} method.

{\bf Returns} {\it connect()} method's respond.

\subsubsection{inet\_sendmsg}
\label{sec:embox_net_inet_sendmsg}
\begin{verbatim}
int inet_sendmsg(struct kiocb *iocb, struct socket *sock,
                 struct msghdr *msg, size_t size)
\end{verbatim}
Send message. Calls protocols native {\it sendmsg()} method.

{\bf Returns} {\it sendmsg()} method's respond.


\subsubsection{inet\_stream\_connect}
\label{sec:af_inet_inet_stream_connect}
\begin{verbatim}
int inet_stream_connect(struct socket *sock, struct sockaddr * addr,
                        int addr_len, int flags)
\end{verbatim}
{\bf Is a wrap} for {\it inet\_dgram\_connect()} for now.

{\bf Returns} the result of {\it inet\_dgram\_connect()} method.

\subsubsection{inet\_listen}
\label{sec:af_inet_inet_listen}
\begin{verbatim}
int inet_listen(struct socket *sock, int backlog)
\end{verbatim}
Calls protocol's native method {\it listen()}.

{\bf Returns} method's respond.

\subsubsection{inet\_accept}
\label{sec:af_inet_inet_accept}
\begin{verbatim}
static int inet_accept(socket_t *sock, sockaddr_t *addr, int *addr_len)
\end{verbatim}
Calls protocol's native method {\it accept()}.

{\bf Returns} method's respond.

\subsection{Structures}
\label{sec:af_inet_structures}
In {\it af\_inet.c} {\it sock} and {\it inet\_sock} structure types are used
widely. This section is the overview of them.
\subsubsection{sock}
Definition is in {\it include/net/sock.h} :
\begin{verbatim}
/**
 * Network layer representation of sockets.
 * @param __sk_common shared layout with inet_timewait_sock
 * @param sk_protocol which protocol this socket belongs in this network family
 * @param sk_type socket type
 * @param sk_rcvbuf size of receive buffer in bytes
 * @param sk_lock synchronizer
 * @param sk_sndbuf size of send buffer in bytes
 * @param sk_flags %SO_LINGER (l_onoff), %SO_BROADCAST, %SO_KEEPALIVE,
 *      %SO_OOBINLINE settings, %SO_TIMESTAMPING settings
 * @param sk_receive_queue incoming packets
 * @param sk_write_queue Packet sending queue
 * @param sk_socket Identd and reporting IO signals
 * @param sk_state_change: callback to indicate change in the state of the sock
 * @param sk_data_ready: callback to indicate there is data to be processed
 * @param sk_write_space: callback to indicate there is bf sending space available
 * @param sk_error_report: callback to indicate errors (e.g. %MSG_ERRQUEUE)
 * @param sk_backlog_rcv: callback to process the backlog
 * @param sk_destruct: called at sock freeing time, i.e. when all refcnt == 0
 */
typedef struct sock {
  struct sock_common __sk_common;
#define sk_family  __sk_common.skc_family
#define sk_prot    __sk_common.skc_prot
#define sk_state   __sk_common.skc_state
  unsigned char sk_protocol;
  unsigned short sk_type;
  int sk_rcvbuf;
  socket_lock_t sk_lock;
  struct {
    struct sk_buff *head;
    struct sk_buff *tail;
  } sk_backlog;
  int sk_sndbuf;
  unsigned long sk_flags;

  struct sk_buff_head *sk_receive_queue;
  struct sk_buff_head *sk_write_queue;

  struct socket *sk_socket;
  void *sk_user_data;

  void (* sk_state_change)(struct sock *sk);
  void (* sk_data_ready)(struct sock *sk, int bytes);
  void (* sk_write_space)(struct sock *sk);
  void (* sk_error_report)(struct sock *sk);
  int (* sk_backlog_rcv)(struct sock *sk, sk_buff_t *pack);
  void (* sk_destruct)(struct sock *sk);
  int (* get_port)(struct sock *sk, unsigned short num);
  int is_ready;
  int answer;
  int sk_err;
} sock_t;

\end{verbatim}

Keeping in mind field {\it sk\_protocol} is vital for development of raw
sockets supporting other higher level protocols. In case of raw sockets
{\it sk\_type} is {\it SOCK\_RAW}.

\subsubsection{inet\_sock}
\label{sec:af_inet_inet_sock}
Definition is in {\it include/net/inet\_sock.c} :
\begin{verbatim}
/**
 * @struct inet_sock
 * @brief Representation of INET sockets
 *
 * @var sk - ancestor class
 * @var pinet6 - pointer to IPv6 control block
 * @var daddr - Foreign IPv4 addr
 * @var rcv_saddr - Bound local IPv4 addr
 * @var dport - Destination port
 * @var num - Local port
 * @var saddr - Sending source
 * @var uc_ttl - Unicast TTL
 * @var sport - Source port
 * @var id - ID counter for DF pkts
 * @var tos - TOS
 * @var mc_ttl - Multicasting TTL
 * @var is_icsk - is this an inet_connection_sock?
 * @var mc_index - Multicast device index
 * @var mc_list - Group array
 * @var cork - info to build ip hdr on each ip frag while socket is corked
 * @var port_type - identificator
 */
typedef struct inet_sock {
  /* sk have to be the first member of inet_sock */
  sock_t         sk;

  in_addr_t      daddr;
  in_addr_t      rcv_saddr;
  struct ip_options *opt;
  __be16         dport;
  __be16         sport;
  __u16          sport_type;
  __u16          num;
  in_addr_t      saddr;
  __s16          uc_ttl;
  __u16          id;
  __u8           tos;
  __u8           mc_ttl;
} inet_sock_t;
\end{verbatim}

%%%%%%%%%%%%% raw_sock.c %%%%%%%%%%%%%%%%%%%%%
\newpage
\section{sk\_buff\_t}
\label{sec:sk_buff_t}
The structure type {\it sk\_buff\_t} is probably widely used across the network
layer of any BSD sockets compatible OS. In EMBOX it has more or less standard
view and is made use of in a similar way. It is described precise enough throught
different papers across internet. It's detailed overview can be simply gained.
Here is how it is defined in EMBOX in {\it include/net/skbuff.h}:

\begin{verbatim}
typedef struct sk_buff {   /* Socket buffer */
  /* These two members must be first. */
  struct sk_buff *next;    /* Next buffer in list */
  struct sk_buff *prev;    /* Previous buffer in list */
  struct sock *sk;         /* Socket we are owned by */
  struct net_device *dev;  /* Device we arrived on/are leaving by */
#if 0
  struct skb_timeval tstamp;  /* Time we arrived */
#endif
  __be16 protocol;         /* Packet protocol from driver */
  uint8_t pkt_type;        /* Packet class */
  char cb[52];             /* Control buffer (used to store layer-specific
                              info e.g. ip options) */
  unsigned int len;        /* Length of actual data */
  union {                  /* Transport layer header */
    struct tcphdr *th;
    struct udphdr *uh;
    struct icmphdr *icmph;
#if 0
    igmphdr *igmph;
    iphdr *ipiph;
    ipv6hdr *ipv6h;
#endif
    unsigned char *raw;
  } h;
  union {                  /* Network layer header */
    struct iphdr *iph;
#if 0
    ipv6hdr *ipv6h;
#endif
    struct arphdr *arph;
    unsigned char *raw;
  } nh;
  union {                  /* Link layer header */
    struct ethhdr *ethh;
    unsigned char *raw;
  } mac;
  __be16 offset;           /* Offset information for ip fragmentation*/
#if 0
  void (*destructor)(struct sk_buff *skb);
#endif
  unsigned char *data;     /* Pointer for buffer used to store all skb content */
  unsigned char *p_data;   /* Pointer for current processing data */
  char prot_info;   /* Protocol level additional data,
                       tcp uses for state handling */
#if 0
  unsigned char tries;
#endif
} sk_buff_t;
\end{verbatim}

In a couple of words it can be described as a container of network information
going in and out of OS. It consists of 3 unions making access to different
protocols' headers' compounds easier. Also it contains bindings to socket to which it
belongs, if there is one, device.

All socket buffers are members of a list, and members {\it next} and {\it prev}
are the pointers to the previous and next list entries.

\newpage
\section{raw\_sock.c}
\label{sec:raw_sock_c}

\subsection{Functions}

\subsubsection{raw\_rcv}
\label{sec:raw_sock_raw_rcv}
\begin{verbatim}
int raw_rcv(sk_buff_t *skb)
\end{verbatim}

{\it raw\_rcv} fetches an {\it sk\_buff\_t} item - an incoming netwrok packet.
This is a standard way for wrapping incoming and outgoing network messages in
EMBOX (See section \ref{sec:sk_buff_t}). Then it cycles thought all raw sockets
opened, which are listed in {\it raw\_hash}(See section \ref{sec:raw_raw_hash}).
It clones incoming skb for each raw socket in {\it raw\_hash}. {\bf TODO}
Here states to realize it without cloning(See section \ref{sec:cloning} on
cloning issue ({\bf TODO create cloning section})).
Then each cloned {\it skb} is being sent to a respected socket.

{\bf Returns} {\it ENOERR} if everything went fine. If something went wrong it
should probably notify about that, but currently it doesn't.

\subsubsection{raw\_err}
\label{sec:raw_sock_raw_err}
\begin{verbatim}
void raw_err(sk_buff_t *skb, uint32_t info)
\end{verbatim}
Notify sockets about error. Cycle though each socket in {\it raw\_has} and set
it's {\it sk\_err} field to {\it info} value.

\subsubsection{raw\_close}
\label{sec:raw_sock_raw_close}
\label{sec:raw_sock_}
\begin{verbatim}
static void raw_close(struct sock *sk, long timeout)
\end{verbatim}
Frees structure {\it sk} given. Probably should perform some more actions.

\subsubsection{raw\_rcv\_skb}
\label{sec:raw_sock_raw_rcv_skb}
\begin{verbatim}
static int raw_rcv_skb(struct sock *sk, sk_buff_t *skb)
\end{verbatim}
Passes given arguments to {\it sock\_queue\_rcv\_skb} method (See section
{\ref{sec:shared_methods}}). It is defined in {\it include/net/sock.h} and
{net/core/sock.c}.

{\bf Returns} {\it NET\_RX\_DROP} if {\it sock\_queue\_rcv\_skb} fails. Else
returns {\it NET\_RX\_SUCCESS}.

\subsubsection{raw\_v4\_hash}
\label{sec:raw_sock_raw_v4_hash}
\begin{verbatim}
static void raw_v4_hash(struct sock *sk)
\end{verbatim}
Adds an entity of type {\it struct sock} to {\it raw\_hash}. The sense is
registration of a socket.

\subsubsection{raw\_v4\_unhash}
\label{sec:raw_sock_raw_v4_unhash}
\begin{verbatim}
static void raw_v4_unhash(struct sock *sk) {
\end{verbatim}
Remove an entity from {\it raw\_hash} list.

\subsubsection{raw\_sendmsg}
\label{sec:raw_sock_raw_sendmsg}
\begin{verbatim}
static int raw_sendmsg(struct kiocb *iocb, struct sock *sk, struct msghdr *msg,
                       size_t len)
\end{verbatim}
Creates an {\it skb} entity of {\it sk\_buff\_t} type. Fills it up and
send to {\it ip\_send\_packet} method of the underlying IP protocol
defined in {\it include/net/ip.h} and {\it net/ipv4/ip\_output.c}.

{\bf Returns} the result of {\it ip\_send\_packet} method.

\subsubsection{raw\_recvmsg}
\label{sec:raw_sock_raw_recvmsg}
\begin{verbatim}
static int raw_recvmsg(struct kiocb *iocb, struct sock *sk, struct msghdr *msg,
                       size_t len, int noblock, int flags, int *addr_len)
\end{verbatim}
Calls {\it skb\_recv\_datagram} method from the sockets shared methods
(See section \ref{sec:shared_methods}). Corrects input {\it len} if needed.
And deallocates orginal {\it skb} in the sense, that this message is fetched
by the end-client.

{\bf Returns} the length of received message.

\subsubsection{raw\_bind}
\label{sec:raw_sock_raw_bind}
\begin{verbatim}
static int raw_bind(struct sock *sk, struct sockaddr *uaddr, int addr_len)
\end{verbatim}
Binds the socket addresses to a given value in {\it uaddr}.

{\bf Returns} 0.

\subsubsection{raw\_setsockopt}
\label{sec:raw_sock_raw_setsockopt}
\begin{verbatim}
static int raw_setsockopt(struct sock *sk, int level, int optname,
                          char *optval, int optlen)
\end{verbatim}
Is a dummy.

{\bf Returns} 0.

\subsubsection{raw\_getsockopt}
\label{sec:raw_sock_raw_getsockopt}
\begin{verbatim}
static int raw_getsockopt(struct sock *sk, int level, int optname,
                          char *optval, int *optlen)
\end{verbatim}
Is a dummy.

{\bf Returns} 0.

\subsubsection{raw\_ioctl}
\label{sec:raw_sock_raw_ioctl}
\begin{verbatim}
static int raw_ioctl(struct sock *sk, int cmd, unsigned long arg)
\end{verbatim}
Is a dummy.

{\bf Returns} 0.

\subsubsection{ip4\_datagram\_connect}
\label{sec:raw_sock_ip4_datagram_connect}
\begin{verbatim}
int ip4_datagram_connect(struct sock *sk, struct sockaddr *uaddr, int addr_len)
\end{verbatim}
Is a dummy.

{\bf Returns} 0.


\subsection{Structures and arrays}
\label{sec:raw_sock_strustures}

\subsubsection{raw\_prot}
\label{sec:raw_sock_structures_raw_prot}
Is a standard entity of type {\it struct proto} to bind protocol methods to
events.

\subsubsection{inet\_sockraw\_ops}
\label{sec:raw_sock_structures_inet_sockraw_ops}
For {\it SOCK\_RAW} sockets; should be the same as {\it inet\_dgram\_ops} but without
{\it udp\_poll}

\subsubsection{raw\_hash}
\label{sec:raw_sock_structures_raw_hash}
This array holds the list of opened raw sockets.

\subsubsection{}
\label{sec:}
\begin{verbatim}
\end{verbatim}

\end{document}
